Magnetic propeties of a two dimentional electron gas strongly coupled to lights

### 2021-04-06 theory/M2DEG.tex
Add forced oscillator solution

    i \hbar \pdv{\psi}{t} = \hat{H}_e(t)\psi

@@

    \text{L.H.S} & = i \hbar \pdv{\psi}{t} =
    i \hbar \bigg( \pdv{\varphi}{t} \phi + \pdv{\phi}{t} \varphi \bigg) =

@@

      \varphi  \pdv{\phi}{t}

@@

    i \hbar\varphi  \pdv{\phi}{t}

@@

    - i \hbar \pdv{t}

@@

>>>>

    i \hbar \dv{\psi}{t} = \hat{H}_e(t)\psi

@@

    \text{L.H.S} & = i \hbar \dv{\psi}{t} =
    i \hbar \bigg( \dv{\varphi}{t} \phi + \dv{\phi}{t} \varphi \bigg) =

@@

      \varphi  \dv{\phi}{t}

@@

    i \hbar\varphi  \dv{\phi}{t}

@@

    - i \hbar \dv{t}
    \phi(y,t) = 0.
If we turn off the external dressing field, this equation leads to simple harmonic oscillator Hamiltonian as follows
\begin{equation} \label{1.30}
    - i \hbar \dv{t}

@@

\begin{equation} \label{1.31}
     i \hbar \dv{\phi(y,t)}{t} =
    \frac{\hat{p}_y^2}{2m_e} +
    \frac{1}{2}m_e \omega_0^2y^2
    \phi(y,t).
Therefore we can identify the $S(t) \equiv eE\sin(\omega t)$ part as a external force act on the harmonic oscillator and we can solve this as a forced harmonic oscillator in $y$ axis.
\begin{equation} \label{1.32}
    i \hbar \dv{\phi(y,t)}{t} =
    \pdv[2]{y} +
    \frac{1}{2}m_e \omega_0^2y^2
    - yS(t)]
    \phi(y,t).
This system can be extacly solvable and we can solve this equation using the methods explained by Husimi [*1] as follows.

\noindent
First we can introduce the time dependent shifted corrdinte as
\begin{equation} \label{1.33}
    y \rightarrow y' = y - \zeta(t) \quad \Rightarrow \quad
    y = y' + \zeta(t)
\end{equation}
and this implies that
\begin{equation} \label{1.34}
    \dv{\phi(y',t)}{t} = \pdv{\phi(y',t)}{t} + \pdv{\phi(y',t)}{y'}\pdv{y'}{t} =
    \pdv{\phi(y',t)}{t} - \dot{\zeta}(t)\pdv{\phi(y',t)}{y'}
\end{equation}
where $\dot{\zeta}(t) = \pdv{\zeta(t)}{t}$.
Therefore, Eq. \eqref{1.32} will be modified to
\begin{equation} \label{1.35}
  \begin{aligned}
    i \hbar \pdv{\phi(y',t)}{t}  =
    \bigg[
    i \hbar\dot{\zeta}\pdv{y'}
    -
    \frac{\hbar^2}{2m_e}
    \pdv[2]{{y'}} +
    \frac{1}{2}m_e \omega_0^2(y' + \zeta)^2
    - (y' + \zeta) S(t)
    \bigg]
    \phi(y',t).
  \end{aligned}
\end{equation}
Let's tranform the wave function using following unitary trasnform
\begin{equation} \label{1.36}
    \phi(y',t) = \exp(\frac{im_e\dot{\zeta}y'}{\hbar})\varphi(y',t)
\end{equation}
and subtitte this into the Eq. \eqref{1.35} and we will get the following
\begin{equation} \label{1.37}
  \text{R.H.S} =  \bigg[ i \hbar \pdv{t} -
i \hbar \Big(\frac{im_e \ddot{\zeta} y'}{\hbar}\Big)\bigg]  \exp(\frac{-im_e\dot{\zeta}y'}{\hbar})\varphi(y',t)
\end{equation}
and
\begin{equation} \label{1.38}
  \begin{aligned}
    \text{L.H.S} & =  \bigg[
      i \hbar \dot{\zeta}  \Big(\frac{im_e \dot{\zeta}}{\hbar}\Big)  +
      i \hbar \dot{\zeta} \pdv{y'} \\
      &
      -
      \frac{\hbar^2}{2m_e}\Big[
        \Big(\frac{im_e \dot{\zeta}}{\hbar}\Big)^2
        + \Big(\frac{2im_e \dot{\zeta}}{\hbar}\Big) \pdv{{y'}}
        + \pdv[2]{{y'}}
      \Big] \\
      &
      +\frac{1}{2}m_e\omega_0^2 {y'}^2 + \frac{1}{2}m_e\omega_0^2 \zeta^2 +
      m_e\omega_0^2 y'\zeta \\
      & -
      y'S(t) - \zeta S(t)
    \bigg]  \exp(\frac{-im_e\dot{\zeta}y'}{\hbar})\varphi(y',t).
  \end{aligned}
\end{equation}
Combining these two we get derive that
\begin{equation} \label{1.39}
  \begin{aligned}
    i \hbar \pdv{\varphi(y',t)}{t}   =
    \bigg[
        -  \frac{\hbar^2}{2m_e}\pdv[2]{{y'}}
        & + \frac{1}{2} m_e \omega_0^2 y'^2 +
        \Big[
            m_e\ddot{\zeta} + m_e\omega_0^2\zeta - S(t)
        \Big]y' \\
        &
        +
        \Big[
            - \frac{1}{2} m_e\dot{\zeta}^2 + \frac{1}{2}m_e\omega_0^2 \zeta^2 - \zeta S(t)
        \Big]
    \bigg]\varphi(y',t).
  \end{aligned}
\end{equation}
Then we can restrict our $\zeta(t)$ function such that
\begin{equation} \label{1.40}
  m_e\ddot{\zeta} + m_e\omega_0^2\zeta = S(t)
\end{equation}
and that leads to
\begin{equation} \label{1.41}
  \begin{aligned}
    i \hbar \pdv{\varphi(y',t)}{t}   =
    \bigg[
        -  \frac{\hbar^2}{2m_e}\pdv[2]{{y'}}
        + \frac{1}{2} m_e \omega_0^2 {y'}^2
        - L(\zeta,\dot{\zeta},t)
    \bigg]\varphi(y',t)
  \end{aligned}
\end{equation}
where
\begin{equation} \label{1.42}
  L(\zeta,\dot{\zeta},t) \equiv \frac{1}{2} m_e\dot{\zeta}^2 - \frac{1}{2}m_e\omega_0^2 \zeta^2 + \zeta S(t)
\end{equation}
is the largrangian of a driven oscillator.

\noindent
Now introduce new unitary transormation for the wavefunction as follows
\begin{equation} \label{1.43}
    \varphi(y',t) = \exp(\frac{i}{\hbar}\int_0^{t}dt'L(\zeta,\dot{\zeta},t')) \chi(y',t)
\end{equation}
and subtite this into the Eq. \eqref{1.41} and gets
\begin{equation} \label{1.44}
  \begin{aligned}
    i \hbar \bigg[
      & \exp(\frac{i}{\hbar}\int_0^{t}dt'L(\zeta,\dot{\zeta},t')) \pdv{t}
      +
      i \hbar L(\zeta,\dot{\zeta},t) \exp(\frac{i}{\hbar}\int_0^{t}dt'L(\zeta,\dot{\zeta},t'))
    \bigg]\chi(y',t) \\
    & =
    \bigg[
        -  \frac{\hbar^2}{2m_e}\pdv[2]{{y'}}
        + \frac{1}{2} m_e \omega_0^2 {y'}^2
        - L(\zeta,\dot{\zeta},t)
    \bigg] \exp(\frac{i}{\hbar}\int_0^{t}dt'L(\zeta,\dot{\zeta},t')) \chi(y',t)
  \end{aligned}
\end{equation}
and finally we can derive that
\begin{equation} \label{1.45}
  \begin{aligned}
    i \hbar \pdv{t} \chi(y',t)  =
    \bigg[
        -  \frac{\hbar^2}{2m_e}\pdv[2]{{y'}}
        + \frac{1}{2} m_e \omega_0^2 {y'}^2
    \bigg] \chi(y',t).
  \end{aligned}
\end{equation}
This is the well known Schrodinger equation of a stationary harmonic oscillator.
In terms of the eigenvalues
\begin{equation} \label{1.46}
  E_n = \hbar \omega_0 \big(n + \frac{1}{2}\big)
\end{equation}
of well-known harmonic eigenfucntions
\begin{equation} \label{1.47}
  \chi_n(y') = \frac{1}{\sqrt{2^{n}n!}}  \cdot
  \bigg(\frac{m_e\omega_0}{\pi \hbar}\bigg)^{1/4}
  \cdot e^{-\frac{m_e\omega_0 y'^2}{2\hbar}} \cdot
  \mathcal{H}_n \bigg(\sqrt{\frac{m_e \omega_0}{\hbar}}y'\bigg)
\end{equation}
being propositional to the Hermite functions $\mathcal{H}_n$, the solutions of Eq. \eqref{1.32} can be represent as
\begin{equation} \label{1.48}
  \phi_n(y,t) = \chi_n(y - \zeta(t))
  \exp(\frac{i}{\hbar}\bigg[- E_nt +
  m_e\dot{\zeta(t)}\big(y-\zeta(t)\big)
   + \int_0^{t}dt'L(\zeta,\dot{\zeta},t')\bigg])
\end{equation}
The set $\chi(y)$ forms a complete set and thus ay general solution $\phi_(y,t)$ can be expaned in terms of the solutions in Eq. \eqref{1.48}.

\noindent
Next we consider special case where we assumed
\begin{equation} \label{1.49}
  S(t) = eE\sin(\omega t)
\end{equation}
and one can derive the Eq. \eqref{1.40} for $\zeta(t)$
\begin{equation} \label{1.50}
  m_e\ddot{\zeta} + m_e\omega_0^2\zeta = eE\sin(\omega t)
\end{equation}
and using Green function method the solution can be write as
\begin{equation} \label{1.51}
  \zeta(t) = \frac{eE}{m_e(\omega_0^2 - \omega^2)}\sin(\omega t).
\end{equation}
form this solution we are able to derive the final solution for the forced harmonic oscillator. \hfill$\blacksquare$
